Optimizing and Incrementalizing Collection Queries by AST Transformations

### 2017-07-19 chap-intro.tex
Intro: more contributions and thesis organization

\Cref{part:incr} makes the following contributions:
\begin{itemize}
\item We present a novel mathematical theory of changes and derivatives, which
  is more general than other work in the field because changes are first-class
  entities, they are distinct from base values and they are defined also for
  functions.
  %(\cref{sec:1st-order-changes}).
  %KO: I think the next sentence cannot be understood at this point.
  %We introduce changes for complex types, defined compositionally.
%
\item We present the first approach to incremental computation for pure
  $\lambda$-calculi by a source-to-source transformation, $\DERIVE$, that
  requires no run-time support. The transformation produces an incremental
  program in the same language; all optimization techniques for the original
  program are applicable to the incremental program as well.
%KO: commented this out. I think the purity is not important enough
%to deserve another sentence here, since we only vaguely hint
%at "further research".
%Since our incremental programs use no impure features, they are
%especially amenable to further optimizations, making this approach
%very suitable for further research.
%
% KO: Let's have one bullet point per section. Also, a conjecture
% sounds like a rather weak contribution
%\item We argue that incrementalization is efficient on
%  \emph{self-maintainable programs}, and discuss how further research on
%  static or dynamic memoization can speed up a larger class of programs (\cref{sec:performance-cons}).
%  \pg{This contribution references text which is now commented
%    out. I believe the text should be brought back in.}
%
\item We prove that our incrementalizing transformation $\DERIVE$
is correct~(\cref{eq:correctness}) by a novel machine-checked logical relation
proof, mechanized in Agda~\citep{agda-head}.

@@

\pg{Describe \cref{part:caching}}
\subsection{Included papers}
This thesis includes material from joint work with colleagues.

@@

\Cref{part:incr} is originally based on work by \citet*{CaiEtAl2014ILC}, though
significantly revised. This work was even more of a team effort. I designed the
overall project and came up with the original notion of change structures; Cai
Yufei contributed differentiation itself and its first correctness proof.
\Cref{part:incr} contains novel correctness proof; its history and
contributions are discussed in \cref{sec:ilc-dev-history}.

@@

>>>>

\Cref{part:incr} makes the following contributions:
\begin{itemize}
\item We present a novel mathematical theory of changes and derivatives, which
  is more general than other work in the field because changes are first-class
  entities, they are distinct from base values and they are defined also for
  functions.
  %(\cref{sec:1st-order-changes}).
  %KO: I think the next sentence cannot be understood at this point.
  %We introduce changes for complex types, defined compositionally.
%
\item We present the first approach to incremental computation for pure
  $\lambda$-calculi by a source-to-source transformation, $\DERIVE$, that
  requires no run-time support. The transformation produces an incremental
  program in the same language; all optimization techniques for the original
  program are applicable to the incremental program as well.

  % \pg{ Maybe comment in? Not sure, which might be telling.}
  % Since our incremental programs use no impure features, they are
  % especially amenable to further optimizations, making this approach
  % very suitable for further research.

%KO: commented this out. I think the purity is not important enough
%to deserve another sentence here, since we only vaguely hint
%at "further research".
% KO: Let's have one bullet point per section. Also, a conjecture
% sounds like a rather weak contribution
%\item We argue that incrementalization is efficient on
%  \emph{self-maintainable programs}, and discuss how further research on
%  static or dynamic memoization can speed up a larger class of programs (\cref{sec:performance-cons}).
%  \pg{This contribution references text which is now commented
%    out. I believe the text should be brought back in.}
%
\item We prove that our incrementalizing transformation $\DERIVE$
is correct~(\cref{eq:correctness}) by a novel machine-checked logical relation
proof, mechanized in Agda~\citep{agda-head}.

@@

\pg{Describe \cref{part:caching}}
\pg{Paste contributions from paper}
\pg{Maybe defunc chapter?}

\pg{WIP}
Other chapters propose further contributions.
\Cref{ch:bsos} proposes the first correctness proofs for ILC via operational
methods and (step-indexed) logical relations, for simply-typed $\lambda$-calculus
(without and with general recursion) and for untyped $\lambda$-calculus.

\Cref{ch:diff-parametricity-system-f} outlines how to extend differentiation to
System F and suggests a proof avenue. Differentiation for System F requires a
generalization of change structures to changes across different types
(\cref{sec:param-derive-changes-across-types}) that appears of independent
interest, though further research remains to be done.

\subsection{Included papers}
This thesis includes material from joint work with colleagues.

@@

\Cref{part:incr} is originally based on work by \citet*{CaiEtAl2014ILC}, though
significantly revised; only \cref{sec:applying} survives mostly unchanged.
This work was even more of a team effort. I designed the
overall project and came up with the original notion of change structures; Cai
Yufei contributed differentiation itself and its first correctness proof.
\Cref{part:incr} contains novel correctness proof; its history and
contributions are discussed in \cref{sec:ilc-dev-history}.

@@

\subsection{Navigating this thesis}
\pg{WIP!!}
Many chapters are independent from each other, with a few exceptions.
\Cref{part:ch-aosd13} can be read independently from the rest.

% Write this even though maybe I restructure the parts.
\Cref{part:incr} and \Cref{part:caching} describe incrementalization.

\paragraph{Differentiation}
\Cref{sec:intro} and \cref{ch:diff-examples} introduce the overall approach to
incrementalization informally.

The core mathematical theory underlying incrementalization correctness using ILC
is developed formally in \cref{ch:derive-formally} and~\cref{ch:change-theory}
using a denotational framework. Equational reasoning on terms is then developed
in~\cref{sec:term-reasoning}. These chapters contain full formal proofs: readers
are welcome to skip or skim those proofs where appropriate. We strive not to
hide important definitions inside theorems, and to give stand-alone definitions
and theorem statements.

Later chapters are again independent from each other.

\paragraph{Performance of differentiated programs}
\Cref{sec:applying} studies how to apply differentiation (as introduced in
previous chapters) to incrementalize a case study and empirically evaluates
performance speedups.

\paragraph{Differentiation with static caching}
\pg{Explain label}
\Cref{ch:cts} is mostly self-contained, even though it builds on the rest of the
material; it summarizes the basics of ILC in \cref{sec:ilc-background}.
Terminology in that chapter is sometimes subtly different from the rest of the thesis.

\pg{At least in more formal chapters, all definitions, notations, and so on are
  clearly highlighted and presented as standalone definitions for ease of reference.}

\pg{Missing chapters}

\paragraph{Numbering}
We number tables and figures per-chapter with a shared counter. For instance,
\cref{fig:measurements-overview} is the first figure in \cref{sec:evaluation},
followed by \cref{table:performance}.

Similarly, we number definitions, remarks, examples and all sorts of theorems
per-chapter with counters including section numbers. For
instance, \cref{def:inc-semantics} is the first definition in
\cref{ch:derive-formally}, followed by \cref{thm:derive-correct}, and they both
appear in \cref{sec:correct-derive}.

All such references are hyperlinked in electronic versions of this thesis.
